### 2023-03-12 article3.qmd
update

The biological carbon pump (BCP) refers to a series of processes that transport biogenic carbon produced in sunlit surface waters to the ocean interior [@Claustre2021-tx]. It is composed of three main pathways [@Claustre2021-tx; @Boyd2019-ez; @Siegel2022-wd]: the biological gravitational pump that is the gravitational settling of particulate organic carbon (POC); the physically mediated particle injection pump that represents physical mechanisms that transport dissolved organic carbon (DOC) to depth via mixing processes; and the biologically mediated particle injection pump that characterizes the transport of POC (and DOC) by diel and seasonal migration of animals in the water column. Historically, the BCP was mainly studied with sediment traps [@Honjo2008-kn] and Thorium-234 radionuclide disequilibria [@Buesseler2007-ql] focused on the gravitational pump hence the contribution of DOC and migrating animals to the BCP was poorly known. However, the advent of autonomous underwater vehicles (AUVs) equipped with a variety of physical and biogeochemical sensors have demonstrated their ability to better understand them [e.g. @Omand2015-mf; @DallOlmo2016-ex; @Llort2018-pa; @Lacour2019-id; @Haentjens2020-dc; @Stukel2017-lc]. Still, we are missing processes to close the carbon budget [@Giering2014-tj; @Boyd2019-ez] and unkowns remain in the under-sampled mesopelagic (i.e. twilight) zone [@Martin2020-wa]. In order to fill the observational gaps, the Biogeochemical-Argo (BGC-Argo) network aims to develop an operational global array of 1000 profiling floats measuring six core variables from the surface to 2000 m [@Claustre2020-jo]: oxygen, nitrate, pH, chlorophyll a (Chla), suspended particles ($b_{bp}$) and downwelling irradiance. In addition to the six core variables, new sensors have been developped to target specific BCP-related questions [@Claustre2020-jo; @Claustre2021-tx; @Chai2020-uq]. In this study, we will focus on BGC-Argo floats equipped with two additional sensors: the Underwater Vision Profiler (UVP6) [@Picheral2021-sx] and a transmissometer used as an optical sediment trap (OST).

The UVP6 is a portable version of the UVP5 [@Picheral2010-to], specifically designed for AUVs [@Picheral2021-sx]. It consists of an underwater camera that counts particles and images objects in, respectively, the 0.1 - 2.5 cm and 0.8 - 2.5 cm size range. In contrast with UVP5 deployments on cabled platforms, UVP6 deployed on long-term drifting AUVs (e.g. BGC-Argo floats) are rarely recovered. Therefore, the latest version of the UVP6 also embarks an embedded image classification algorithm that transmits the abundance of 20 classes (16 biological classes) in near-real time. The OST consists of a vertically mounted transmissometer whose upward-facing optical window intercepts sinking particles. The latter have already been used on AUVs [@Bishop2004-at; @Bishop2009-ka; @Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk] and have demonstrated their usefulness to characterize the gravitational pump.

For the first time, two pairs of BGC-Argo floats equipped with both the UVP6 and an OST were deployed (May 2022) in the Labrador Sea (LS) to investigate the seasonal and episodic dynamics of particle and carbon fluxes. The LS is a basin of the subpolar North Atlantic (NA) ocean which represents 20% of the global ocean carbon sink \[@Takahashi2009-rv\]. The efficiency of this sink is explained by both deep winter convection that traps CO$_{2}$ from the atmosphere \[@Vage2008-qa; @Yashayaev2009-fe; @Yashayaev2017-mn\] and by the largest spring phytoplankton bloom \[@Siegel2002-hk\] that consumes CO$_{2}$ for the photosynthetic production of organic matter. The triggering mechanism of the NA bloom was initially explained by the 'Critical Depth Hypothesis' of @Sverdrup1953-ki that defines a critical surface mixing depth where phytoplankton growth is equal to phytoplankton losses. According to @Sverdrup1953-ki, a bloom is therefore triggered when the surface mixed layer depth (MLD) is shallower than this critical depth resulting in a positive phytoplankton growth rate. However, this hypothesis was challenged by @Behrenfeld2010-my who suggested that the initiation of the NA blooms begins in winter with a slow phytoplankton biomass increase over the following months (i.e. 'Dilution-Recoupling' hypothesis). These hypotheses have been investigated with BGC-Argo floats but the debate still remains \[@Mignot2018-gm; @Yang2020-hw; @Boss2010-do\]. For the LS, @Lacour2015-da used a climatology of satellite Chla to divide the NA into 6 bioregions with specific bloom phenologies. In their bioregionalization (see also their Figure 2), the LS is characterized by two bioregions with distinct bloom phenologies on both side of the 60°N parallel. The northern part is characterized by a short spring bloom in late April whereas the bloom in the southern part starts in April-May and slowly develops to reach its maximum in June. @Lacour2015-da concluded that the concurrent influence of surface light and mixing was the first-order mechanism controlling the onset of the spring bloom in both bioregions. This is coherent with the results of @Wu2008-vw and @Frajka-Williams2010-ph who found that early shoaling of the mixed layer was responsible for the early spring bloom in the northern part whereas the light controlled the bloom in the southern part.

In late May 2022, two pairs of BGC-Argo floats were deployed in the southern part of the LS at the time of the spring bloom. While BGC-Argo floats typically drift at 1000 m for \~10 days between each ascending profile, these two pairs of floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study POC and particle fluxes derived from beam attenuance \[@Estapa2023-oz\] and UVP6 particle size distributions (PSD). The goals of this study are threefold: 1) Compare the UVP6 and OST as two independent measurements of downward carbon fluxes and particle abundance; 2) Explore the seasonal signal of the downward POC flux and particle abundance with Chla and MLD using BGC-Argo floats and satellite data; 3) Evaluate the new drifting configuration and how it could be improved in regards with our preliminary results.

@@

Two couples of BGC-Argo floats were deployed in the Labrador Sea in late May 2022. The first couple (WMO 6904240 and 4903634) was deployed on the 22$^{nd}$ at coordinates (-49.3°E, 58.9°N) while the second (WMO 6904241 and 1902578) was deployed on the 30$^{th}$ at coordinates (-52.4°E, 56.8°N).

All 4 floats used in this study are PROVOR Jumbo floats (NKE Instrumentation, France) equipped with a CTD SBE41-CP (Sea-Bird Scientific, USA), a UVP6-LP (Hydroptic, France), a c-Rover 660 nm beam transmissometer (WETLabs, USA) and a ECO FLBBCD (WETLabs, USA) that includes a chlorophyll-a (Chla) fluorometer (excitation at 470 nm, emission at 695 nm), a colored dissolved organic matter (CDOM) fluorometer (excitation at 370 nm, emission at 460 nm) and a 700 nm particle backscatteringmeter ($b_{bp}$). All floats but one (1902578) were equipped with a OCR-504 Multispectral Radiometer to measure the photosynthetic available radiation.

All 4 floats have been configured to drift at 3 distinct parking depths (respectively, 200 m, 500 m and 1000 m for approximately 1, 3 and 5 days) to study the downward carbon flux. During these drift phases, measurements of particles abundance (14 size classes, between 0.1 mm and 2.5 mm) are taken every 20 minutes at 200 m and every 2h at 500 m and 1000 m whereas $c_{p}$ measurements are taken every 30 minutes at all depths. This configuration can be adjusted during the mission of the float via two-way Iridium communication.

@@

Data were downloaded from the Coriolis data center (*ftp://ftp.ifremer.fr/ifremer/argo/*, with biogeochemical data at *dac/coriolis/* and UVP6 data at *aux/coriolis/*, last accessed: March 6, 2023). Chla data were quality controlled for Non-Photochemical Quenching (NPQ), a photoprotective mechanism that decreases the apparent fluorescence per unit of Chla following the method implemented by @Terrats2020-ef who modified the method of @Xing2018-pn for shallow-mixing cases. Chla data were also divided by a slope factor of 2.15 following the recommendation of @Roesler2017-me for obtaining unbiased Chla estimates from in situ WETLabs ECO fluorometer. $c_{p}$ data were converted from counts to physical units (m$^{-1}$) using the following relationship

@@

UVP6 data and in particular the averaged number of particles per size class (i.e. particle size distribution hereafter referred to as PSD) was divided by the number of images at each depth and the water volume imaged by the UVP (0.7L) to obtain particle concentrations (units: #/L). Note that in an impending update by the Coriolis data center, it should not be necessary to divide by the number of images anymore.

@@

In order to compute $F_{small}$, we firstly computed the slope of each group of observations between jumps (positive and negative, see @fig-estapa-method_article). Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window (e.g. turbulence ou zooplankton consumption of accumulated material). Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each remaining group in order to derive the small particles flux using the linear relationship between the beam attenuance flux and the POC flux (see first panel of Figure 2 in @Estapa2023-oz).

@@

```{r, example of a cp signal analyzed for jumps (with an old method here)}
#| label: fig-estapa-method_article
#| fig-cap: Application of the method of @Estapa2017-jr on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component and a discontinuous (positive jumps, red) component. Slopes for each group are depicted in light blue. A negative jump was also detected (yellow). Data from BGC-ARGO float WMO 6904240, drifting at 1000 m during its 21$^{st}$ cycle.

@@

ggplot(tmp2, aes(x = dates, y = cp, colour = colour)) + geom_point(size = 3) +
    scale_colour_manual(values = c('#003366','#E31B23', '#FFC325')) + 
    #scale_shape_manual(values = c(1,1,4)) +
    geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = '#DCEEF3', size = 2) + theme_bw() + 
    labs(x = 'Time', y = TeX('$c_{p}~(m^{-1})$')) +
  geom_point(data = check, aes(x = dates, y = cp, colour = colour), size = 3) +
  geom_point(data = check, aes(x = dates, y = cp), shape = 1, size = 3, colour = 'black')  +
    theme(legend.position = 'none')
```

@@

Satellite Chla data were downloaded from a daily-merged cloud free GlobColour product (doi: [10.48670/moi-00281](https://doi.org/10.48670/moi-00281)) with a spatial resolution of 4 km. Mixed layer depth (MLD) derived from satellite observations were downloaded from near-real-time weekly data from the ARMOR3D product (doi: [10.48670/moi-00052](https://doi.org/10.48670/moi-00052)).

@@

$F_{small}$ and $F_{big}$ were computed at each drifting depth for all floats (@fig-OST-flux). At 200 m, $F_{small}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ (average $\pm$ standard deviation). At 500 m, $F_{small}$ is comprised between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_smallf, 1)` mg C m$^{2}$ day$^{-1}$. For both drifting depths, the trend shows a decrease of $F_{small}$ from June to, respectively, January and February. Note that the drifting configuration of all 4 floats was adapted to the deepening of the MLD (Fig. XX) hence drifting at 200 m and 500 m was stopped at this period to avoid measurements bias due to turbulence. At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid/end of July followed by a steady decrease towards March, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.

```{r, plot Fsmall and Fbig, fig.width=8, fig.height=6}
#| label: fig-OST-flux
#| fig-cap: OST-derived small (top) and big (bottom) particles carbon flux for all floats at each drifting depth. Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean. There is no data after January and February at, respectively, 200 m and 500 m because the drifting was stopped at those depths due to the increase of the MLD.

@@

up / down

@@

In the upper part of the size range, @fig-uvp6-profiling-512-2500 shows a rapid emptying of the entire water column starting in November. In contrast with the small classes, some particles, especially above 1.5 mm, have already virtually disappeared from the water column by September. However, the rapid increase and decrease of particles on the entire water column around July-August for floats 6904241 and 1902578 (still in the same zone at that time) could suggest an export event.

@@

### Environmental conditions before and shortly after deployment

All floats measured similar particle concentrations at each drifting depth and on the vertical. In addition, except perhaps for the upper part of the UVP6 size spectrum for floats 6904241 and 1902578, the results do not indicate a clear hint of particle export signature. Therefore, we suggest that the 4 BGC-Argo floats mostly measured the dynamics of the seasonal, continuous, particle signal but that they missed the main transient particle export events associated with the LS spring bloom.

To estimate the time of the bloom in the deployment area, we compared MLD and surface (0 - 20 m) Chla concentrations from both satellite and in situ data (@fig-chla-mld-satellite) in a 100 km radius around each deployment. It can be seen that these estimates are in agreement, therefore we hypothesize that we can trust satellite data prior to the floats deployment. Taking this into consideration, it seems that the pair 6904240-4903634 missed a bloom that occurred in early May whereas the pair 6904241-1902578 missed a first Chla increase in late May days before their deployment. However, a second equivalent increase (median of 2 mg m$^{-3}$) happened in late June - early July, which could be a contributing factor to the apparent particle track observed around July and August.

```{r, chla and MLD from satellite, fig.height=6, fig.width=6}
#|: label: fig-chla-mld-satellite
#| fig-cap: Comparison of MLD and Chla concentrations measured from satellites (black) and BGC-Argo floats (red) following their deployment. Satellite error bars represent the first and third quartiles around the median in a 50 km radius around each deployment (each pair of floats) coordinates. In situ Chla represent the Chla median in the upper 20 m of the water column with the upper and lower incertitude in the correction factor of @Roesler2017-me.

@@

@@

chla_cover_deployment2 <- points_in_circle(chla_sat, lon_center = -52.3, lat_center = 56.7, radius = 10000) # 50 km around center
chla_cover_deployment2 <- chla_cover_deployment2 |> dplyr::mutate(date = as_date(time), group = '6904241 - 1902578') |> dplyr::mutate(deploy = as_date('2022-05-30'))

@@

chla_plot <- ggplot(chla_cover |> filter(date > '2022-03-01')) + geom_errorbar(aes(x = date, y = chla),
                      stat = 'summary',
                      fun.min = function(z) {quantile(z, 0.25)},
                      fun.max = function(z) {quantile(z, 0.75)},
                      fun = median, colour = 'black', width = .5) + theme_bw() +
  labs(x = 'Month', y = TeX('Chla (mg m$^{-3}$)'))+ facet_wrap(~group) +
  geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "black", linetype = 'dashed') +
  geom_pointrange(data = d_chla |> filter(date < '2022-08-01'), aes(x = date, ymin = median_chla_0_20m_down, ymax = median_chla_0_20m_up, y = median_chla_0_20m), colour = "#E31B23", size = .25) +
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month')

>>>>

The biological carbon pump (BCP) refers to a series of processes that transport biogenic carbon produced in sunlit surface waters to the ocean interior [@Claustre2021-tx]. It is composed of three main pathways [@Claustre2021-tx; @Boyd2019-ez; @Siegel2022-wd]: the biological gravitational pump that is the gravitational settling of particulate organic carbon (POC); the physically mediated particle injection pump that represents physical mechanisms that transport dissolved organic carbon (DOC) to depth via mixing processes; and the biologically mediated particle injection pump that characterizes the transport of POC (and DOC) by diel and seasonal migration of animals in the water column. Historically, the BCP was mainly studied with sediment traps [@Honjo2008-kn] and Thorium-234 radionuclide disequilibria [@Buesseler2007-ql] focused on the gravitational pump hence the contribution of DOC and migrating animals to the BCP was poorly known. However, the advent of autonomous underwater vehicles (AUVs) equipped with a variety of physical and biogeochemical sensors have demonstrated their ability to better understand them [e.g. @Omand2015-mf; @DallOlmo2016-ex; @Llort2018-pa; @Lacour2019-id; @Haentjens2020-dc; @Stukel2017-lc]. Still, we are missing processes to close the carbon budget [@Giering2014-tj; @Boyd2019-ez] and unkowns remain in the under-sampled mesopelagic (i.e. twilight) zone [@Martin2020-wa]. In order to fill the observational gaps, the Biogeochemical-Argo (BGC-Argo) network aims to develop an operational global array of 1000 profiling floats measuring six core variables from the surface to 2000 m [@Claustre2020-jo]: oxygen, nitrate, pH, chlorophyll a (Chla), suspended particles and downwelling irradiance. In addition to the six core variables, new sensors have been developped to target specific BCP-related questions [@Claustre2020-jo; @Claustre2021-tx; @Chai2020-uq]. In this study, we will focus on BGC-Argo floats equipped with two additional sensors: the Underwater Vision Profiler (UVP6) [@Picheral2021-sx] and a transmissometer used as an optical sediment trap (OST).

The UVP6 is a portable version of the UVP5 [@Picheral2010-to], specifically designed for AUVs [@Picheral2021-sx]. It consists of an underwater camera that counts particles and images objects in, respectively, the 0.1 - 2.5 cm and 0.8 - 2.5 cm size range. In contrast with UVP5 deployments on cabled platforms, UVP6 deployed on long-term drifting AUVs (e.g. BGC-Argo floats) are rarely recovered. Therefore, the latest version of the UVP6 also embarks an embedded image classification algorithm that transmits the abundance of 20 classes (16 biological classes) in near-real time. The OST consists of a vertically mounted transmissometer whose upward-facing optical window intercepts sinking particles to estimate the downward POC flux. The latter have already been used on AUVs [@Bishop2004-at; @Bishop2009-ka; @Estapa2013-gw; @Estapa2017-jr; @Estapa2019-nk] and have demonstrated their usefulness to characterize the gravitational pump.

For the first time, BGC-Argo floats equipped with both the UVP6 and an OST were deployed in the Labrador Sea (LS) to investigate the seasonal and episodic dynamics of particles and carbon fluxes. The LS is a basin of the subpolar North Atlantic (NA) ocean which represents 20% of the global ocean carbon sink [@Takahashi2009-rv]. The efficiency of this sink is explained by both deep winter convection that traps CO$_{2}$ from the atmosphere [@Vage2008-qa; @Yashayaev2009-fe; @Yashayaev2017-mn] and by the largest spring phytoplankton bloom [@Siegel2002-hk] that consumes CO$_{2}$ for the photosynthetic production of organic matter. The triggering mechanism of the NA bloom was initially explained by the 'Critical Depth Hypothesis' of @Sverdrup1953-ki that defines a critical surface mixing depth where phytoplankton growth is equal to phytoplankton losses. According to @Sverdrup1953-ki, a bloom is therefore triggered when the surface mixed layer depth (MLD) is shallower than this critical depth resulting in a positive phytoplankton growth rate. However, this hypothesis was challenged by @Behrenfeld2010-my who suggested that the initiation of the NA blooms begins in winter with a slow phytoplankton biomass increase over the following months (i.e. 'Dilution-Recoupling' hypothesis). These hypotheses have been investigated with BGC-Argo floats but the debate still remains [@Mignot2018-gm; @Yang2020-hw; @Boss2010-do]. For the LS, @Lacour2015-da used a climatology of satellite Chla to divide the NA into 6 bioregions with specific bloom phenologies. In their bioregionalization (see also their Figure 2), the LS is characterized by two bioregions with distinct bloom phenologies on both side of the 60°N parallel. The northern part is characterized by a short spring bloom in late April whereas the bloom in the southern part starts in April-May and slowly develops to reach its maximum in June. @Lacour2015-da concluded that the concurrent influence of surface light and mixing was the first-order mechanism controlling the onset of the spring bloom in both bioregions. This is coherent with the results of @Wu2008-vw and @Frajka-Williams2010-ph who found that early shoaling of the mixed layer was responsible for the early spring bloom in the northern part whereas the light controlled the bloom in the southern part. In late May 2022, two pairs of BGC-Argo floats were deployed in the southern part of the LS at the time of the spring bloom. While BGC-Argo floats typically drift at 1000 m for 10 days between each ascending profile, all floats were configured to drift at three distinct parking depths (200 m, 500 m and 1000 m) to study POC fluxes and particle concentrations derived from beam attenuance [@Estapa2023-oz] and UVP6 particle size distributions (PSD).

The goals of this study are threefold: 1) the comparison of the UVP6 and OST as two independent measurements of downward carbon fluxes and particle abundance on BGC-Argo floats; 2) the exploration of the seasonality (bloom and non-bloom periods) of downward POC fluxes and particle abundance in the LS; 3) the evaluation of the new drifting configuration and how it could be improved to study specific BCP-related questions.

@@

Two couples of BGC-Argo floats were deployed in the Labrador Sea in late May 2022. The first couple (WMO 6904240 and 4903634) was deployed on the 22$^{nd}$ at coordinates (-49.3°E, 58.9°N) while the second (WMO 6904241 and 1902578) was deployed on the 29$^{th}$ at coordinates (-52.4°E, 56.8°N).

All 4 floats used in this study are PROVOR Jumbo floats (NKE Instrumentation, France) equipped with a CTD SBE41-CP (Sea-Bird Scientific, USA), a UVP6-LP (Hydroptic, France), a c-Rover 660 nm beam transmissometer (WETLabs, USA) and a ECO FLBBCD (WETLabs, USA) that includes a Chla fluorometer (excitation at 470 nm, emission at 695 nm), a colored dissolved organic matter (CDOM) fluorometer (excitation at 370 nm, emission at 460 nm) and a 700 nm particle backscatteringmeter ($b_{bp}$). All floats but one (1902578) were equipped with a OCR-504 Multispectral Radiometer to measure the photosynthetic available radiation.

All 4 floats have been configured to drift at 3 distinct parking depths (respectively, 200 m, 500 m and 1000 m for approximately 1, 3 and 5 days) to study the downward carbon flux. During these drift phases, measurements of particle abundance (14 size classes, between 0.1 mm and 2.5 mm) are taken every 20 minutes at 200 m and every 2h at 500 m and 1000 m whereas $c_{p}$ measurements are taken every 30 minutes at all depths. This configuration can be adjusted during the mission of the float via two-way Iridium communication.

@@

Data were downloaded from the Coriolis data center (*ftp://ftp.ifremer.fr/ifremer/argo/*, with biogeochemical data at *dac/coriolis/* and UVP6 data at *aux/coriolis/*, last accessed: March 6, 2023). Chla data were quality controlled for Non-Photochemical Quenching (NPQ), a photoprotective mechanism that decreases the apparent fluorescence per unit of Chla following the method implemented by @Terrats2020-ef who modified the method of @Xing2018-pn for shallow-mixing cases. Chla data were also divided by a slope factor of 2.15 following the recommendation of @Roesler2017-me for obtaining unbiased Chla estimates from in situ WETLabs ECO fluorometers. Temperature and salinity data were quality controlled using @Wong2022-hz. The potential density anomaly was computed with the function swSigma0 of the oce R package [@Kelley_undated-sw] and the MLD was determined using the 0.03 kg m$^{-3}$ density criterion at a reference depth of 10 m [@De_Boyer_Montegut2004-hh]. $c_{p}$ data were converted from counts to physical units (m$^{-1}$) using the following relationship

@@

UVP6 data and in particular the averaged number of particles per size class (i.e. PSD) was divided by the number of images at each depth and the water volume imaged by the UVP (0.7L) to obtain particle concentrations (units: #/L). Note that in an impending update by the Coriolis data center, it will no longer be necessary to divide by the number of images.

@@

In order to compute $F_{small}$, we firstly computed the slope of each group of observations between jumps (positive and negative, see @fig-estapa-method_article). Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by factors other than the pump cleaning the optical window (e.g. turbulence ou zooplankton consumption of accumulated material). Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each remaining group in order to derive the small particles flux using the linear relationship between the beam attenuance flux and the POC flux (see first panel of Figure 2 in @Estapa2023-oz).

@@

```{r, example of a cp signal analyzed for jumps (with an old method here), fig.width=8, fig.height=5}
#| label: fig-estapa-method_article
#| fig-cap: Application of the method of @Estapa2017-jr on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component and a discontinuous (positive jumps, red) component. Slopes for each group are depicted in light blue. A negative jump was also detected (yellow). Data from BGC-ARGO float WMO 6904240, drifting at 1000 m during its 21$^{st}$ cycle.

@@

ggplot(tmp2, aes(x = dates, y = cp, colour = colour)) + geom_point(size = 3) +
    scale_colour_manual(values = c('#003366','#E31B23', '#FFC325')) + 
    #scale_shape_manual(values = c(1,1,4)) +
    geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = '#DCEEF3', linewidth = 1) + theme_bw() + 
    labs(x = 'Time', y = TeX('$c_{p}~(m^{-1})$')) +
  geom_point(data = check, aes(x = dates, y = cp, colour = colour), size = 3) +
  geom_point(data = check, aes(x = dates, y = cp), shape = 1, size = 3, colour = 'black')  +
    theme(legend.position = 'none') + scale_x_datetime(breaks = c(as.POSIXct('2022-11-16'), as.POSIXct('2022-11-18')), date_labels = 'Nov %d')
```

@@

Satellite Chla data were downloaded from a daily-merged cloud free GlobColour product (doi: 10.48670/moi-00281) with a spatial resolution of 4 km. Mixed layer depth (MLD) derived from satellite observations were downloaded from near-real time weekly data from the ARMOR3D product (doi: 10.48670/moi-00052).

@@

$F_{small}$ and $F_{big}$ were computed at each drifting depth for all floats (@fig-OST-flux). At 200 m, $F_{small}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ (average $\pm$ standard deviation). At 500 m, $F_{small}$ is comprised between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_smallf, 1)` mg C m$^{2}$ day$^{-1}$. For both drifting depths, the trend shows a decrease of $F_{small}$ from June to, respectively, January and February before the drift was stopped at those depths due to the deepening of the winter MLD (@fig-mld) in order to ensure the float stability. At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid/end of July followed by a steady decrease towards March, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.

```{r, Fsmall and Fbig, fig.width=8, fig.height=6}
#| label: fig-OST-flux
#| fig-cap: OST-derived small (top) and big (bottom) particles carbon flux for all floats at each drifting depth. Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean. There is no data after January and February at, respectively, 200 m and 500 m because the drifting was stopped at those depths due to the deepening of the MLD.

@@

up / down
```

```{r, MLD, fig.height=4, fig.width=8}
#| label: fig-mld
#| fig-cap: MLD measured by all floats from the end of May 2022 to early March 2023. Black dashed lines represent the first two drifting depths.

# add MLD here
mld_data <- vroom('/home/flo/dox/ArgoShine/npq_corrected_data.csv')
# clean and smoothed chla data
mld_data <- mld_data |> select(cycle, MLD, pres, juld, wmo, chla_npq) |> dplyr::mutate(date = as_date(juld), WMO = factor(wmo, levels = c(6904240,4903634,6904241,1902578)))
mld_data <- mld_data |> group_by(date, WMO, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))

# reverselog_trans <- function(base = exp(1)) {
#     trans <- function(x) -log(x, base)
#     inv <- function(x) base^(-x)
#     trans_new(paste0("reverselog-", format(base)), trans, inv, 
#               log_breaks(base = base), 
#               domain = c(1e-100, Inf))
# }
# 
# mld_plot <- ggplot(mld_data) + geom_line(aes(x = date, y = mld, colour = WMO)) +
#   labs(x = 'Month', y = 'MLD (m)') + scale_y_continuous(trans = reverselog_trans(10), breaks = c(10,50,200,400,750)) +
#   scale_color_brewer(palette = 'Dark2') + 
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + theme_bw()

# reverselog_trans <- function(base = exp(1)) {
#     trans <- function(x) -log(x, base)
#     inv <- function(x) base^(-x)
#     trans_new(paste0("reverselog-", format(base)), trans, inv, 
#               log_breaks(base = base), 
#               domain = c(1e-100, Inf))
# }

mld_plot <- ggplot(mld_data) + geom_line(aes(x = date, y = mld, colour = WMO)) +
  labs(x = 'Month', y = 'MLD (m)') + scale_y_reverse() +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') + theme_bw() +
  geom_hline(yintercept = 200, colour = 'black', linetype = 'dashed') +
  geom_hline(yintercept = 500, colour = 'black', linetype = 'dashed')

mld_plot
```

@@

In the upper part of the size range, @fig-uvp6-profiling-512-2500 shows a rapid emptying of the entire water column starting in November. In contrast with small classes (referred to here as below 512 $\mu$m), some particles, especially above 1.5 mm, have almost entirely disappeared from the water column by September. However, the rapid increase and decrease of particles on the entire water column around July-August for floats 6904241 and 1902578 (still in the same zone at that time) could suggest an export event.

@@

### Phenology of the spring bloom in the Labrador Sea

<!-- All floats measured similar particle concentrations at each drifting depth and on the vertical during the entire time series. In addition, except perhaps for the upper part of the UVP6 size spectrum for floats 6904241 and 1902578, the results do not indicate a clear hint of particle export signature. Therefore, we suggest that the 4 BGC-Argo floats mostly measured the dynamics of the seasonal, continuous, particle signal but that they missed the main transient particle export events associated with the LS spring bloom. -->

To estimate the time of the bloom in the deployment area, we compared MLD and surface (i.e. upper 20 m) Chla concentrations from both satellite and in situ data in a 100 km radius around each deployment. @fig-chla-mld-satellite shows that satellite and in situ Chla and MLD data concur therefore we assume that the former can be used to study environmental conditions before the floats deployment. Taking this into consideration, @fig-chla-mld-satellite indicates that the 6904240-4903634 pair missed a bloom that occurred in early May whereas the 6904241-1902578 pair missed a first (satellite) Chla increase at the end of May days before their deployment. However, a second (satellite and in situ) Chla increase reached a peak at the end of June before it decreased again towards July.

```{r, chla and MLD from satellite, fig.height=6, fig.width=6}
#| label: fig-chla-mld-satellite
#| fig-cap: Comparison of MLD and Chla concentrations measured from satellites (dark blue) and BGC-Argo floats (red) following their deployment (in pairs). Satellite error bars represent the first and third quartiles around the median in a 100 km radius around each deployment coordinates. In situ Chla represents the Chla median in the upper 20 m of the water column with uncertainties from the slope factor uncertainty of @Roesler2017-me.

@@

# remove NaN in smoothed chla
d <- d |> drop_na(smoothed_chla)

@@

chla_cover_deployment2 <- points_in_circle(chla_sat, lon_center = -52.3, lat_center = 56.7, radius = 10000) # 50 km around center
chla_cover_deployment2 <- chla_cover_deployment2 |> dplyr::mutate(date = as_date(time), group = '6904241 - 1902578') |> dplyr::mutate(deploy = as_date('2022-05-29'))

@@

# HPLC data
# chla_hplc <- vroom('DATA/ARGO/CE22009_pigments_031122_Travail_Herve_FŔ.csv') |> filter(Station %in% c(29,9)) |> select(depth = `Depth_(m)`, chla = Total_Chlorophyll_a, station = Station) |> filter(depth < 10) |> mutate(group = if_else(station == 9, '6904240 - 4903634', '6904241 - 1902578')

chla_plot <- ggplot(chla_cover |> filter(date > '2022-03-01')) + geom_errorbar(aes(x = date, y = chla),
                      stat = 'summary',
                      fun.min = function(z) {quantile(z, 0.25)},
                      fun.max = function(z) {quantile(z, 0.75)},
                      fun = median, colour = '#003366', width = .5) + theme_bw() +
  labs(x = 'Month', y = TeX('Chla (mg m$^{-3}$)'))+ facet_wrap(~group) +
  geom_vline(data = chla_cover, aes(xintercept = deploy), colour = "black", linetype = 'dashed') +
  geom_pointrange(data = d_chla |> filter(date < '2022-08-01'), aes(x = date, ymin = median_chla_0_20m_down, ymax = median_chla_0_20m_up, y = median_chla_0_20m), colour = "#E31B23", size = .25) +
   scale_x_date(labels = date_format("%m"), date_breaks = '1 month')
